\section*{Acknowledgment}

I want to express my sincere gratitude to my supervisor from CERN, \textbf{Maria Dimou}. Over the last nine months, I have learned beyond the content of this thesis; I would be incapable of putting the insights from these attainments into writing. I have enjoyed our weekly supervision meetings, which had often drifted from the planned agenda into exchanges over life beyond our solar system, the philosophy of ancient Greeks, intuitionism in mathematics, challenges in learning the German language, political decisions in tumultuous times, and much more. Thank you for teaching me so much; I will miss the weekly chat undoubtedly, but I am sure it is no goodbye.

\vspace{1cm}

I had the great pleasure to work with \textbf{Michiel De Jong} for the longevity of this project. He has given me every needed resource to understand Solid and be part of the community. His deep understanding of the Web, political values, and immense motivation to build purposeful software for everyone is admirable and has become a role model to me.

\vspace{1cm}

\textbf{Adrian M\"onnich}, chief developer of Indico at CERN, provided me with answers to every possible question I could have asked in the context of Indico, technicalities at CERN, or general advice in Web engineering. Thank you for listening and giving me all the information I needed for this project.

\vspace{1cm}

And one more expression of gratitude towards \textbf{Sir Tim Berners-Lee} for giving input on the possible scope developing the prototypes, showing curiosity towards my implementations, and providing feedback several times. It has been a great honor working so closely and on a project as impactful as Solid.
