\chapter{Conclusion}
% \setcounter{section}{0}

Decentralizing a centralized system is complex, and even developing and integrating small-scale Solid modules that harmonize with sophisticated software is. The in this paper shown work has shown in an experimental approach how the Solid ecosystem can enrich \gls{cern} software, its challenges and has laid out a possible future for the collaboration of the two.

The paper presented two working Solid apps embedded into Indico and has shown how data can reside in user's control while still be served purposeful to others. The applications have also shown how existing information in one's data pod can be used in other systems. These two insights from storing data outside of Indico and enabling information to flow back into the system has shown admirable features for the future of the Web: control through external data pods and reuse thanks to interoperable data formats of Linked Data. The third principle of Solid, decentralized authentication was also enabled and necessary for the development, but had a lesser impact due to restricting the prototype to only Indico and not guiding testing participants to other applications. Further, a decentralized \gls{sso} service is already in use at \gls{cern}.

The implementation of Solid comes with imperfections. The detailed analysis and evaluation of the Solid apps have shown those flaws in the synergy between Indico and Solid principles. The flaws being mainly performance regressions in the retrieval and distribution of data from several external data pods. Where in centralized systems the data is stored in one place and can be processed and prepared for clients neatly, in a decentralized architecture several requests are required to retrieve the data first. The prototypes' design was as decoupled as possible from Indico to not infiltrate Indico's functionality and allow other systems to integrate the modules as well. The decoupling has proven that the two systems are too distant and implementations need to be closer to Indico. A more imminent solution means Indico is required to change too. The data flow from the data pod back into Indico, as demonstrated in the second implementation of the \gls{poc}, is challenging by reason of lacking any semantic structure whatsoever in Indico.

The \glspl{poc} and analyses have displayed that decentralization brings many challenges for the development of applications. However, as Tim Berners-Lee once said: “Data is a precious thing and will last longer than the systems themselves.” it is sensible to decouple applications from data to allow innovation not through owning data, but by building applications without relying on having the data.

\vspace{0.5cm}

Solid as an ecosystem is active, maturing, and should be closely followed. Startups are forming worldwide and innovate by developing applications for the Solid ecosystem. Governments in several countries show their interest by investing and funding Solid-related projects to enable its citizens data pods and opportunity for data control. 

The ecosystem is still inexperienced due to the scarcity of available applications for end-users to use. An atrocious \gls{ui} for the currently most popular data pod implementation hinders newcomers from testing Solid for private use. Whereas the \gls{ui} of the implementation can be fixed by using more apparent interactions with the system, Solid still has many usability challenges to account for. Applications and agents need workflows for sharing data so seamless and without being in doubt for a second when handling their sensitive personal information. The limited number of applications and developers actively participating makes Solid still very experimental and theoretical. A planned \textit{call-for-implementation} by the Solid technical writers will encounter these challenges and find solutions for the interesting performance complications found in the \gls{poc} development. There is no doubt that in the following years, the Solid implementations and their use will skyrocket. However, it will take more organizations, companies, and individuals to initiate a breakthrough until then. The products developed need to be clever, and their benefits need to be visible to the end-user. When the end-user cannot distinguish a decentralized from a centralized solution, and no frictions are visible on use, Solid will be a success.

% * Solid lives through the community, when only one company does the implementation it will not thrive
% * vicious cycle
% * 

% Solid from the community, for everyone
% start by the community
% get going, but breakthrough takes more -> CERN

% community <-> company
% not just one company it won't thrive

% software needs to be rethought
% major challenges:
%     usability
%     performance
%     describing the world in Linked Data, can the data of the complex world be described for by machines?
% it takes entities as CERN to not let commercial companies built the products, otherwise Solids futures might be at stake? (a bit harsh)

% solid is still experimental and needs a lot of attention from developers
% startups are good
% the more the merrier
% the community needs to remain as a driver
% high profile people help, but cannot be the motivator

% build better and hybrid products for people where they might not even realize the decentralized benefits

% challenge the known